Changes in accessibility to emergency and community food services during COVID-19 and implications for low income populations in Hamilton, Ontario

### 2021-10-13 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.Rmd
Updated README after publication in Social Science and Medicine

For this study, we aim to look at how the landscape of emergency food and related services (e.g. low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic. Did the number of open services diminish? If so, what was the accessibility to emergency and community food services before the pandemic from the perspective of low income households, and how has it changed during the pandemic with respect to geographic access and congestion at remaining sites? And finally, who are most likely to have been impacted by changes in the accessibility landscape? This paper first looks at the distribution of emergency and community food services before and during the pandemic. Then, we use the balanced floating catchment area approach of Paez et al. [-@paez2019demand] to investigate the accessibility situation. For this, we adopt a fully disaggregated approach based on parcel-level data. Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016. This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open]. The data and code needed to reproduce the analysis are available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing).

@@

Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in an (anonymous) [Drive Folder](https://drive.google.com/drive/folders/1-l8hO1pMIqaBqf57j_M_BMrXhXFvrV2c?usp=sharing). The contents of the data package are described next.

>>>>

For this study, we aim to look at how the landscape of emergency food and related services (e.g. low-cost or free meal service providers) available in Hamilton, Ontario, changed during the pandemic. Did the number of open services diminish? If so, what was the accessibility to emergency and community food services before the pandemic from the perspective of low income households, and how has it changed during the pandemic with respect to geographic access and congestion at remaining sites? And finally, who are most likely to have been impacted by changes in the accessibility landscape? This paper first looks at the distribution of emergency and community food services before and during the pandemic. Then, we use the balanced floating catchment area approach of Paez et al. [-@paez2019demand] to investigate the accessibility situation. For this, we adopt a fully disaggregated approach based on parcel-level data. Socio-economic and demographic data are drawn from the latest Census of Canada (2016), whereas travel information is from the most recent regional travel survey from 2016. This paper follows reproducible research recommendations [see @brunsdon2020opening], and the research was conducted using open source tools for transportation analysis [@lovelace2021open]. The data and code needed to reproduce the analysis are available in a [public repository](https://github.com/paezha/Accessibility-Food-Banks-Hamilton).

@@

Data have been prepared for sharing in the form of an open data product [see @arribas2021open] available in a [public repository](https://github.com/paezha/Accessibility-Food-Banks-Hamilton). The contents of the data package are described next.
